---
title: "Coursera: getting and cleaning data"
author: "Jessica Grindheim"
date: '`r Sys.Date()`'
output:
  html_document:
    fig_width: 7
    fig_height: 6
    fig_caption: true
    toc: yes
    toc_float: true
    toc_depth: 6
    df_print: paged
  pdf_document: 
    toc: yes
    toc_depth: 6
latex_engine: xelatex
geometry: left = 1in, right = 1in, top = 1in, bottom = 1in
---

```{r setup , include=FALSE}
knitr::opts_chunk$set(
    echo = F
  , message = F
  , warning = F
  , fig.fullwidth = T
  , fig.align = "center"
  , fig.margin = T
  , tidy.opts = list(width.cutoff=80)
  ) 

# the following colors the boxes in knit html
# https://www.r-bloggers.com/2017/06/pretty-errors-warnings-and-messages-in-r-markdown/
knitr::knit_hooks$set(
   error = function(x, options) {
     paste('\n\n<div class="alert alert-danger">',
           gsub('##', '\n', gsub('^##\ Error', '**Error**', x)),
           '</div>', sep = '\n')
   },
   warning = function(x, options) {
     paste('\n\n<div class="alert alert-warning">',
           gsub('##', '\n', gsub('^##\ Warning:', '**Warning**', x)),
           '</div>', sep = '\n')
   },
   message = function(x, options) {
     paste('\n\n<div class="alert alert-info">',
           gsub('##', '\n', x),
           '</div>', sep = '\n')
   }
)
```


# Background

* **Goals**  
  * Take notes during the course  
  * Do necessary exercises  
  



```{r, include = F}
library("here")
source(here::here("scripts","003.getting_and_cleaning_data","003.config.R"))

```

# Module 1- Raw and processed data

* **download.file**  
  * dateDownloaded = date()   
  * if https on mac, use method = "curl"  
* **reading xml**  
  * exensible markup language  
  * frequently used to store structured data  
  * XML::xmlTreeParse    
    * rootNode = XML::xmlRoot() to get root node  
    * XML::xpathSapply( rootNode, xmlValue)  # nice to grab all the values out  
* **reading JSON**  
  * package: jsonlite  
  
* **data.table**  
  * faster and more efficient that data.table  
  * inherits from data.frame, so functions that you use on data.frames should still work  
  * but requires learning of new syntax  
  
## dplyr
*  dplyr supplies five 'verbs' that cover most fundamental data manipulation tasks: select(), filter(), arrange(), mutate(), and summarize()  
  * cran is tibble I'm playing with in a swirl session    

  
    
### select (for columns)

```{r}
# select columns
# to select only the ip_id, package, and country variables from the cran dataset.  
select(cran, ip_id, package, country) 

# select columns with : 
# to select all columns starting from r_arch and ending with country.  
select(cran, r_arch:country) 

# omit columns
# to omit the time column  
select(cran, -time) 
  
```

### filter (for rows)

  
```{r}
# filter rows based on column value 
filter(cran, package == "swirl")    
# filter rows based on 2 column values
filter(cran, r_version == "3.1.1", country== "US")   
# filter rows based on x column OR y column
filter(cran, country == "US" | country == "IN")  
# filter nas with is.na() function**  

```

### arrange (to order rows)  


```{r}
# sort ascending order
  arrange(cran2, ip_id)  
# sort descending order
  arrange(cran2, desc(ip_id) ) 
# sort multiple variables
  arrange(cran2, country, desc(r_version), ip_id)  
```

### mutate to create a new variable  

```{r}
# create a new column
mutate(cran3, size_mb = size / 2^20)  

# create 2 new columns, one based off of the first column!
mutate(cran3, size_mb = size / 2^20, size_gb = size_mb / 2^10 )  

```

### summarize and group_by
**summarize** collapses the dataset into a single row. More powerful when applied to grouped data.    

```{r}
# summarize a column
summarize(cran, avg_bytes = mean(size))  

# summarize a column after grouping by another column
summarize(cran, avg_bytes = mean(size))  

# group by 1 variable
by_package <- group_by(cran, package)  

# summarize an already grouped tibble (not sure it has to be a tibble to work)
summarize( by_package, mean(size))  
```

```{r}


# Compute four values, in the following order, from
# the grouped data:
#
# 1. count = n()
# 2. unique = n_distinct(ip_id)
# 3. countries = n_distinct(country)
# 4. avg_bytes = mean(size)
#
# A few thing to be careful of:
#
# 1. Separate arguments by commas
# 2. Make sure you have a closing parenthesis
# 3. Check your spelling!
# 4. Store the result in pack_sum (for 'package summary')
#
# You should also take a look at ?n and ?n_distinct, so
# that you really understand what is going on.

pack_sum <- summarize(by_package
                      , count = n()
                      , unique = dplyr::n_distinct(ip_id) 
                      , countries = dplyr::n_distinct(country)
                      , avg_bytes = mean(size) 
                      )

```

  
    
    
    
    
    


# Module 2

# Module 3


# Module 4# Session info

```{r}

sessionInfo()
knitr::knit_exit()
```


